Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-12-31 manuscript.Rmd
Fix formatting for the jounal

The first level of cohesion that we will consider are the words within social
media posts.  In Figure \@ref(fig:cohesion-in-tweets), we see four sequential
tweets of four words each. Each sequential tweet contains some content included
in the previous tweets. The linguistic construct of content word overlap
measures the relationship across units of meaning---here, tweets---where
identical and semantically similar words are repeated [@graesser2004coh]. These
repeated lexical items constitute a theme; on Twitter, this might be a
particular hashtag, a name, a location, or another nominal category. This
localized cohesion is referential; words are repeated and co-referred across
tweets, creating a sense of continuity and cohesiveness.

[Figure \@ref(fig:cohesion-in-tweets) about here]

@@

### Sociolinguistic Cohesion: Connectedness of Speakers, Ideas, and Events
We tie the constructs of lexical, interlocutor, and social cohesion together in
a theoretical framework called *sociolinguistic cohesion*, as shown in Figure
\@ref(fig:theoretical-framework). Previous values of linguistic cohesion should
influence future values of linguistic cohesion and future actions in the real
world space. Previous events influence future events and future linguistic
cohesion in social media.

[Figure \@ref(fig:theoretical-framework) about here]

@@

This linguistic cohesion metric is similar to *term frequency-inverse document
frequency* [tf-idf; see also @jones1972statistical;
@robertson2004understanding]. However, the method focuses more squarely on terms
(i.e., agnostic to the number of documents in the corpus), aggregates at a
higher semantic level (i.e., summing across words), and accounts for linguistic
novelty relative to the pair and the corpus (i.e., ratio of Eq. 1 to Eq. 2). The
social cohesion calculation was performed on the tweet corpus containing English
tweets. All hashtags were included in the analysis, while URLs were removed. The
raw social cohesion time series can be seen in Figure
\@ref(fig:raw-cohesion-ts).

[Figure \@ref(fig:raw-cohesion-ts) about here]

@@

The raw event time series for target filtered data can be seen in Figure
\@ref(fig:raw-event-ts-target). The raw event time series for source and target
filtered data can be seen in Figure \@ref(fig:raw-event-ts-source-target).

[Figures \@ref(fig:raw-event-ts-target) and
\@ref(fig:raw-event-ts-source-target) about here]

@@

To perform categorical CRQA, the time series must be converted to the same
scale, while still preserving a sampling rate that respects the native timescale
of the system [cf. Nyquist frequency in signal processing;
@grenander1959probability]. In other words, a system that changes every day must
be sampled more often than a system that changes every year and less often than
a system that changes every hour. However, to compare two time series with CRQA,
the two time series must be sampled at the same frequency as one another.
Because the social cohesion and event data have different scales and values, we
converted each time series of daily values (e.g., Figure \@ref(fig:raw-ts)) into
deciles (e.g., Figure \@ref(fig:deciled-ts)) to analyze the relative dynamics of
different levels of social cohesion and counts of all, positive, and negative
events on each day. We implemented CRQA using the `crqa` package [@R-crqa] for R
[@R-base].

[Figures \@ref(fig:raw-ts) and \@ref(fig:deciled-ts) about here]

@@

We can visually analyze these plots to reveal patterns of structure and
recurrence. The main diagonal line on the plot is known as the *line of
synchrony*. This line falls where the time series match temporally. A completely
filled main diagonal line would indicate perfect in-phase synchrony between the
two time series. See Figures
\@ref(fig:plot-rp-targ)-\@ref(fig:plot-rp-source-targ) for examples of RPs.

@@

#### Visualization
To visualize WCRQA, the resulting metrics from each window can be plotted as a
time series. This allows for visual interpretation of the pattern or fluctuation
of the metrics across time. These dynamics can then be linked to other external
events. See Figures \@ref(fig:plot-wcrqa-targ-all) and
\@ref(fig:plot-wcrqa-source-targ-all) for examples.

@@

A permutation test for the CRQA of social cohesion and count of all events found
a statistically significant difference from chance for DET, NRLINE, and LAM. All
other metrics did not reach significance (Figure \@ref(fig:plot-rp-targ);
\autoref{table-1}).

A permutation test for the CRQA of social cohesion and count of positive events
found a statistically significant difference from chance for DET, NRLINE, maxL,
LAM, and TT. All other metrics did not reach significance (Figure
\@ref(fig:plot-rp-targ); \autoref{table-1}).

A permutation test for the CRQA of social cohesion and count of negative events
found a statistically significant difference from chance for DET, NRLINE, and
maxL. All other metrics did not reach significance (Figure
\@ref(fig:plot-rp-targ); \autoref{table-1}).

[Figure \@ref(fig:plot-rp-targ) and \autoref{table-1} about here]

@@

### Syria as Source and Target
A permutation test for the CRQA of social cohesion and count of all events found
a statistically significant difference from chance for DET, maxL, and LAM. All
other metrics were non-significant (Figure \@ref(fig:plot-rp-source-targ);
\autoref{table-2}).

A permutation test for the CRQA of social cohesion and count of positive events
found a statistically significant difference from chance for DET, L, ENTR,
rENTR, and LAM (Figure \@ref(fig:plot-rp-source-targ); \autoref{table-2}).

A permutation test for the CRQA of social cohesion and count of negative events
found no statistically significant difference from chance for and metrics
(Figure \@ref(fig:plot-rp-source-targ); \autoref{table-2}).

[Figure \@ref(fig:plot-rp-source-targ) and \autoref{table-2} about here]

@@

In the WCRQA plots (see Figures \@ref(fig:plot-wcrqa-targ-all) and
\@ref(fig:plot-wcrqa-source-targ-all)), the 95th percentiles are represented by
horizontal red lines, and the 99th percentiles are represented by horizontal
orange lines. The trend lines are plotted in blue.

@@

For social cohesion and count of all events, there was a consistent period of at
or above-chance RR from the 3rd--12th and 44th--50th window. DET was largely
above chance from the 47th to the 54th window, with some exceptions (Figure
\@ref(fig:plot-wcrqa-targ-all)).

For social cohesion and count of positive events, we saw a downward trend in RR
as the window advances throughout the sample. Interestingly, we see largely
above-chance RR until the 16th window, with an exception for window 1. A peak
also appeared for the 40th--42nd windows. DET was intermittently at or above
chance for the 8th--14th and 54th--62nd windows (Figure
\@ref(fig:plot-wcrqa-targ-all)).

For social cohesion and count of negative events, RR stayed above or almost
above chance from the 5th--6th and 46th--49th windows and was below chance for
an extended period from the 56th--62nd window. DET bottomed out at zero for the
15th--23rd windows. The metric was above or nearly above chance only toward the
end of the time series, in the 47th--54th and 62nd windows (Figure
\@ref(fig:plot-wcrqa-targ-all)).

[Figure \@ref(fig:plot-wcrqa-targ-all) about here]

@@

For social cohesion and count of all events, there was a consistent period of
above- or nearly above-chance RR in the first 23 windows and the 44th--52nd
windows. RR was consistently below chance for the 29th--39th windows. DET showed
above- or nearly above-chance values in windows 7--11 and 59--62 (Figure
\@ref(fig:plot-wcrqa-source-targ-all)).

For social cohesion and count of positive events, there was a downward trend in
RR as the window moved, with above- or nearly above-chance values in the first
23 windows and the 57th--62nd windows. RRR was largely below-chance for the
30th--38th windows. DET stayed largely within chance with intermittent
bottoming-out to a value of 0 (Figure \@ref(fig:plot-wcrqa-source-targ-all)).

For social cohesion and negative events, RR was above chance for the 7th window
and below chance for the 15th--19th windows, intermittently below chance
beetween the 25th--39th windows, and below chance for the 51st--62nd windows.
DET showed a downward trend as the windows moved, with no abovee-chance values.
It bottomed out to zero for the 15th--20th and 24th-27th windows, intermittently
bottomed out for the 34th--45th windows, and consistently bottomed out for the
52nd--62nd windows. (Figure \@ref(fig:plot-wcrqa-source-targ-all)).

[Figure \@ref(fig:plot-wcrqa-source-targ-all) about here]

@@

### Syria as Target
For the RR values in WCRQA of social cohesion and count of all events (Figure
\@ref(fig:plot-wcrqa-targ-all)), we see both an early and late above-chance
peak. However, DET for these periods only rises above chance for the later peak.
During this later period, the modal ICEWS event for each day (Figure
\@ref(fig:plot-mode-event)) was almost consistently -10, the most violent event
category. This supports our first hypothesis, that overall online cohesion may
reflect the frequency of events occurring on the ground, particularly those that
are violent in nature.

[Figure \@ref(fig:plot-mode-event) about here]

For positive events, we see a similar early peak in RR, but the later
above-chance values did not match up temporally with that of the all-events time
series (Figure \@ref(fig:plot-wcrqa-targ-all)). This suggests that the early
stretch of above-chance RR may have been captured in the RR of the all-events
analysis, whereas the shorter stretches were overshadowed by the coupling with
negative events. The early period of above-chance DET in the positive-event
analysis aligned with the early period of above-chance RR, whereas the later
stretch of significant DET values aligns temporally with that of the all-events
analysis. It is possible that the later stretch reflected the decrease in the
number of positive events happening (see Figure \@ref(fig:raw-event-ts-target)),
while the relative sparsity of positive events led to an increase in coupling,
with denizens rallying around dwindling bits of hope as time went on.

>>>>

The first level of cohesion that we will consider are the words within social
media posts.  In Fig \@ref(fig:cohesion-in-tweets), we see four sequential
tweets of four words each. Each sequential tweet contains some content included
in the previous tweets. The linguistic construct of content word overlap
measures the relationship across units of meaning---here, tweets---where
identical and semantically similar words are repeated [@graesser2004coh]. These
repeated lexical items constitute a theme; on Twitter, this might be a
particular hashtag, a name, a location, or another nominal category. This
localized cohesion is referential; words are repeated and co-referred across
tweets, creating a sense of continuity and cohesiveness.

[Fig \@ref(fig:cohesion-in-tweets) about here]

@@

### Sociolinguistic Cohesion: Connectedness of Speakers, Ideas, and Events
We tie the constructs of lexical, interlocutor, and social cohesion together in
a theoretical framework called *sociolinguistic cohesion*, as shown in Fig
\@ref(fig:theoretical-framework). Previous values of linguistic cohesion should
influence future values of linguistic cohesion and future actions in the real
world space. Previous events influence future events and future linguistic
cohesion in social media.

[Fig \@ref(fig:theoretical-framework) about here]

@@

This linguistic cohesion metric is similar to *term frequency-inverse document
frequency* [tf-idf; see also @jones1972statistical;
@robertson2004understanding]. However, the method focuses more squarely on terms
(i.e., agnostic to the number of documents in the corpus), aggregates at a
higher semantic level (i.e., summing across words), and accounts for linguistic
novelty relative to the pair and the corpus (i.e., ratio of Eq. 1 to Eq. 2). The
social cohesion calculation was performed on the tweet corpus containing English
tweets. All hashtags were included in the analysis, while URLs were removed. The
raw social cohesion time series can be seen in Fig
\@ref(fig:raw-cohesion-ts).

[Fig \@ref(fig:raw-cohesion-ts) about here]

@@

The raw event time series for target filtered data can be seen in Fig
\@ref(fig:raw-event-ts-target). The raw event time series for source and target
filtered data can be seen in Fig \@ref(fig:raw-event-ts-source-target).

[Figs \@ref(fig:raw-event-ts-target) and
\@ref(fig:raw-event-ts-source-target) about here]

@@

To perform categorical CRQA, the time series must be converted to the same
scale, while still preserving a sampling rate that respects the native timescale
of the system [cf. Nyquist frequency in signal processing;
@grenander1959probability]. In other words, a system that changes every day must
be sampled more often than a system that changes every year and less often than
a system that changes every hour. However, to compare two time series with CRQA,
the two time series must be sampled at the same frequency as one another.
Because the social cohesion and event data have different scales and values, we
converted each time series of daily values (e.g., Fig \@ref(fig:raw-ts)) into
deciles (e.g., Fig \@ref(fig:deciled-ts)) to analyze the relative dynamics of
different levels of social cohesion and counts of all, positive, and negative
events on each day. We implemented CRQA using the `crqa` package [@R-crqa] for R
[@R-base].

[Figs \@ref(fig:raw-ts) and \@ref(fig:deciled-ts) about here]

@@

We can visually analyze these plots to reveal patterns of structure and
recurrence. The main diagonal line on the plot is known as the *line of
synchrony*. This line falls where the time series match temporally. A completely
filled main diagonal line would indicate perfect in-phase synchrony between the
two time series. See Figs
\@ref(fig:plot-rp-targ)-\@ref(fig:plot-rp-source-targ) for examples of RPs.

@@

#### Visualization
To visualize WCRQA, the resulting metrics from each window can be plotted as a
time series. This allows for visual interpretation of the pattern or fluctuation
of the metrics across time. These dynamics can then be linked to other external
events. See Figs \@ref(fig:plot-wcrqa-targ-all) and
\@ref(fig:plot-wcrqa-source-targ-all) for examples.

@@

A permutation test for the CRQA of social cohesion and count of all events found
a statistically significant difference from chance for DET, NRLINE, and LAM. All
other metrics did not reach significance (Fig \@ref(fig:plot-rp-targ);
\autoref{table-1}).

A permutation test for the CRQA of social cohesion and count of positive events
found a statistically significant difference from chance for DET, NRLINE, maxL,
LAM, and TT. All other metrics did not reach significance (Fig
\@ref(fig:plot-rp-targ); \autoref{table-1}).

A permutation test for the CRQA of social cohesion and count of negative events
found a statistically significant difference from chance for DET, NRLINE, and
maxL. All other metrics did not reach significance (Fig
\@ref(fig:plot-rp-targ); \autoref{table-1}).

[Fig \@ref(fig:plot-rp-targ) and \autoref{table-1} about here]

@@

### Syria as Source and Target
A permutation test for the CRQA of social cohesion and count of all events found
a statistically significant difference from chance for DET, maxL, and LAM. All
other metrics were non-significant (Fig \@ref(fig:plot-rp-source-targ);
\autoref{table-2}).

A permutation test for the CRQA of social cohesion and count of positive events
found a statistically significant difference from chance for DET, L, ENTR,
rENTR, and LAM (Fig \@ref(fig:plot-rp-source-targ); \autoref{table-2}).

A permutation test for the CRQA of social cohesion and count of negative events
found no statistically significant difference from chance for and metrics
(Fig \@ref(fig:plot-rp-source-targ); \autoref{table-2}).

[Fig \@ref(fig:plot-rp-source-targ) and \autoref{table-2} about here]

@@

In the WCRQA plots (see Figs \@ref(fig:plot-wcrqa-targ-all) and
\@ref(fig:plot-wcrqa-source-targ-all)), the 95th percentiles are represented by
horizontal red lines, and the 99th percentiles are represented by horizontal
orange lines. The trend lines are plotted in blue.

@@

For social cohesion and count of all events, there was a consistent period of at
or above-chance RR from the 3rd--12th and 44th--50th window. DET was largely
above chance from the 47th to the 54th window, with some exceptions (Fig
\@ref(fig:plot-wcrqa-targ-all)).

For social cohesion and count of positive events, we saw a downward trend in RR
as the window advances throughout the sample. Interestingly, we see largely
above-chance RR until the 16th window, with an exception for window 1. A peak
also appeared for the 40th--42nd windows. DET was intermittently at or above
chance for the 8th--14th and 54th--62nd windows (Fig
\@ref(fig:plot-wcrqa-targ-all)).

For social cohesion and count of negative events, RR stayed above or almost
above chance from the 5th--6th and 46th--49th windows and was below chance for
an extended period from the 56th--62nd window. DET bottomed out at zero for the
15th--23rd windows. The metric was above or nearly above chance only toward the
end of the time series, in the 47th--54th and 62nd windows (Fig
\@ref(fig:plot-wcrqa-targ-all)).

[Fig \@ref(fig:plot-wcrqa-targ-all) about here]

@@

For social cohesion and count of all events, there was a consistent period of
above- or nearly above-chance RR in the first 23 windows and the 44th--52nd
windows. RR was consistently below chance for the 29th--39th windows. DET showed
above- or nearly above-chance values in windows 7--11 and 59--62 (Fig
\@ref(fig:plot-wcrqa-source-targ-all)).

For social cohesion and count of positive events, there was a downward trend in
RR as the window moved, with above- or nearly above-chance values in the first
23 windows and the 57th--62nd windows. RRR was largely below-chance for the
30th--38th windows. DET stayed largely within chance with intermittent
bottoming-out to a value of 0 (Fig \@ref(fig:plot-wcrqa-source-targ-all)).

For social cohesion and negative events, RR was above chance for the 7th window
and below chance for the 15th--19th windows, intermittently below chance
beetween the 25th--39th windows, and below chance for the 51st--62nd windows.
DET showed a downward trend as the windows moved, with no abovee-chance values.
It bottomed out to zero for the 15th--20th and 24th-27th windows, intermittently
bottomed out for the 34th--45th windows, and consistently bottomed out for the
52nd--62nd windows. (Fig \@ref(fig:plot-wcrqa-source-targ-all)).

[Fig \@ref(fig:plot-wcrqa-source-targ-all) about here]

@@

### Syria as Target
For the RR values in WCRQA of social cohesion and count of all events (Fig
\@ref(fig:plot-wcrqa-targ-all)), we see both an early and late above-chance
peak. However, DET for these periods only rises above chance for the later peak.
During this later period, the modal ICEWS event for each day (Fig
\@ref(fig:plot-mode-event)) was almost consistently -10, the most violent event
category. This supports our first hypothesis, that overall online cohesion may
reflect the frequency of events occurring on the ground, particularly those that
are violent in nature.

[Fig \@ref(fig:plot-mode-event) about here]

For positive events, we see a similar early peak in RR, but the later
above-chance values did not match up temporally with that of the all-events time
series (Fig \@ref(fig:plot-wcrqa-targ-all)). This suggests that the early
stretch of above-chance RR may have been captured in the RR of the all-events
analysis, whereas the shorter stretches were overshadowed by the coupling with
negative events. The early period of above-chance DET in the positive-event
analysis aligned with the early period of above-chance RR, whereas the later
stretch of significant DET values aligns temporally with that of the all-events
analysis. It is possible that the later stretch reflected the decrease in the
number of positive events happening (see Fig \@ref(fig:raw-event-ts-target)),
while the relative sparsity of positive events led to an increase in coupling,
with denizens rallying around dwindling bits of hope as time went on.
